---
title: 'Figure 3: Underlying skin microbiome (fungal and bacterial communities) integrated
  with C. auris colonization status'
author: "Diana Proctor"
date: "06/30/2020"
output:
  html_document: default
  pdf_document: default
---
last updated: `r format(Sys.Date(), format="%B %d %Y")`


***

**draft: submission to Nature Medicine October 2020**


*Manuscript Title*: Integrated genomic, epidemiologic investigation of Candida auris skin colonization 

*Authors*: Diana M Proctor1, Thelma Dangana2, D. Joseph Sexton3, Christine Fukuda2, Rachel D Yelin2, Mary Stanley2, Pamela B Bell2, Sangeetha Baskaran2, Clay Deming1, Qiong Chen1, Sean Conlan1, NISC Comparative Sequencing Program4, Rory M Welsh3, Snigdha Vallabhaneni3,5, Tom Chiller3, Kathleen Forsberg3, Stephanie R. Black6, Massimo Pacilli6, Heidi H Kong7, Michael Y Lin2, Michael E Schoeny8, Anastasia P Litvintseva3, Julia A Segre1+, Mary K Hayden2+


*Affiliations*:

- 1Microbial Genomics Section, Translational and Functional Genomics Branch, National Human Genome Research Institute, National Institutes of Health, Bethesda, MD 20892, USA

- 2 Department of Internal Medicine, Division of Infectious Diseases, Rush University Medical Center, Chicago, IL 60612, USA.

- 3Mycotic Diseases Branch, Centers for Disease Control and Prevention, Atlanta, GA 30333, USA.

- 4NIH Intramural Sequencing Center, National Human Genome Research Institute, National Institutes of Health, Bethesda, MD 20892, USA.

- 5Division of Healthcare Quality Promotion, NCEZID, CDC, USA

- 6Communicable Disease Program, Chicago Department of Public Health, Chicago, IL, 60612, USA.

- 7Dermatology Branch, National Institute of Arthritis and Musculoskeletal and Skin Diseases, National Institutes of Health, Bethesda, MD 20892, USA

- 8College of Nursing, Rush University, Chicago, IL 60612, USA.

- +Contributed equally

***

### Description of the dataset

For this data set, we have samples for 51 subjects who were sampled for 16S rRNA and ITS1 sequencing. In addition, we have a clinical data set including the variables that were used to generate Table 1. The purpose of this script is to identify associations between bacteria and fungi with respect to Candida auris colonization outcomes (0, 1).

In order to accomplish this, we read in the following data:

1. its_match.rds
2. bac_match.rds


### Here, we render the following figures:

1. Figure 2
2. Supplementary Figure 5

---

```{r}
#set global knitting options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
```


```{r}
packages <- c("knitr", 
              "tidyverse",
              "colorspace",
              "viridis", 
              "scales", 
              "kableExtra", 
              "ggplot2", 
              "reshape2", 
              "gridExtra", 
              "phyloseq",
              "ggordiplots",
              "DESeq2",
              "mixOmics",
              "ggpubr",
              "dabestr",
              "compositions",
              "ggsci",
              "RColorBrewer",
              "ggrepel",
              "stringr",
              "yarrr",
              "mixOmics")

# install packages from bioconductor
BiocManager::install(setdiff(packages,installed.packages()), update=FALSE)

#From Dan Sprockett
# load packages
n <- length(packages) - sum(sapply(packages,require,character.only=TRUE))

# print if packages loaded properly
if(n == 0){
  print("All necessary R packages loaded properly")
} else {
  print(paste0(n, " R packages did not load properly"))
}


```

set plotting options and define functions

```{r}
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
mypal = brewer.pal(9, "Set1")

ISU_secondary_palette <- c("#3E4827", "#76881D", "#A2A569",
                           "#003D4C", "#006BA6", "#7A99AC",
                           "#7C2529", "#9A3324", "#BE531C",
                           "#8B5B29", "#B9975B", "#EED484",
                           "#6E6259", "#707372", "#ACA39A", "#C8102E")



phyToDf <- function(phy, level) {
      ra = transform_sample_counts(phy, function(x) x/sum(x))
      #domain.phy <- tax_glom(ra, taxrank=level)
      tax.count <- data.frame(data.frame(ra@tax_table@.Data, t(otu_table(ra))))
      dfm = melt(tax.count, colnames(tax_table(ra))) 
      colnames(dfm)[colnames(dfm) == 'variable'] <- 'Fungal.Matcher'
      sample_data(ra)$Fungal.Matcher =
        str_replace_all(sample_data(ra)$Fungal.Matcher, "([;])", ".")
      df = plyr::join(dfm,data.frame(sample_data(ra)) )
      return(df)
}


#define some functions
add_NSubjects <- function(phy) {
    map = as.data.frame(as.matrix(sample_data(phy))) %>%
      dplyr::count(., PID, sort = FALSE, name = "NSamplesPerSubject")  %>%
      plyr::join(., data.frame(sample_data(phy)), by = "PID") %>%
      tibble::column_to_rownames(., "Fungal.Matcher")%>%
      sample_data(.)
}

add_NSites <- function(phy) {
      map = as.data.frame(as.matrix(sample_data(phy))) %>%
      dplyr::group_by(., Unique_ptid) %>%
      dplyr::count(., SiteID,  name = "NSamplesPerSubjectSite")
      df = as.data.frame(map)
      new_map = merge(df, sample_data(phy), by=c("Unique_ptid", "SiteID"))
      rownames(new_map) = new_map$Fungal.Matcher
      new_map = sample_data(new_map)
      return(new_map)
}
```

Read in in the data and clean it up

rename the candida genera to candida
rename the highest rank to the candida variant of the species name

g__Candida s__jadinii = Candida utils
Issatchenkia orientalis = Candida krusei

one asv assigned to the genus candida which couldn't be assigned to the species level and was preesent in only 59 reads, dropped
```{r}
its_match = readRDS(file="~/Desktop/candida_auris_rush/its_match_CDI-out.rds") %>%
                  subset_samples(., Survey_Period==1) %>%
  subset_samples(., Unique_ptid !=32)
sample_names(its_match) = str_replace_all(sample_names(its_match), "Bu/To", "BuTo")
sample_data(its_match)$Fungal.Matcher = sample_names(its_match)

tax = data.frame(its_match@tax_table@.Data) 
taxa_names(its_match) = tax$Seq
write.csv(tax, "~/Desktop/candida_auris_rush/its_match_CDI-out-tax_uncorrected.csv")

tax = read.csv("~/Desktop/candida_auris_rush/its_match_CDI-out-tax_corrected.csv")
rownames(tax) = tax$X
tax = as.matrix(tax) %>%
  tax_table(.)
tax_table(its_match) = tax

#update taxa names to highest rank
tax = data.frame(its_match@tax_table@.Data)
new.names = tax$Highest.Rank
taxa_names(its_match) = new.names


#drop taxa that aren't assigned to the genus level

its_match2 = subset_taxa(its_match, !(ASV_Number %in% c(
            "ASV_458",
            "ASV_647",
            "ASV_1314",
            "ASV_1400",
            "ASV_1724",
            "ASV_1792",
            "ASV_1860",
            "ASV_1861",
            "ASV_1897",
            "ASV_2340",
            "ASV_3824")))

its_match2 =  subset_taxa(its_match2, Highest.Rank != "Less_than_10_per_ASV Less_than_10_per_ASV")
#read in the bacterial data
bac_match = readRDS(file="~/Desktop/candida_auris_rush/bac_match_CDI-out.rds") %>%
  subset_samples(., Unique_ptid !=32) %>%
  subset_samples(., Survey_Period=="1") %>%
  prune_taxa(taxa_sums(.) > 0, .)

sample_names(bac_match) = str_replace_all(sample_names(bac_match), "Bu/To", "BuTo")
sample_data(bac_match)$Fungal.Matcher = sample_names(bac_match)

filtergroup = genefilter::filterfun(genefilter::kOverA(k=76, A=10)) #k = number of samples; A = abundance
bac_filt =filter_taxa(bac_match, filtergroup, prune=TRUE)  %>%
        prune_taxa(taxa_sums(.) > 0, .) 


its_filt = its_match2


### get the clinical data
clindf = data.frame(sample_data(its_filt),
                    Unique_ptid=sample_data(its_filt)$Unique_ptid,
                    Site=sample_data(its_filt)$site,
                    Cauris_Result = sample_data(its_filt)$Cauris_Result,
                    Survey_Period=sample_data(its_filt)$Survey_Period)
 clindf = dplyr::select(clindf, c("Unique_ptid", "Survey_Period", "Site",  "Cauris_Result",
                     "trach", "gtube", "urinary_cath", "mech_vent", "age", "braden_score",
                     "pasthospitaladmit", "antifungal_rx", "abx_rx", "contact_iso",
                     "sex", "braden_score"))

clindf$Cauris_Result = as.factor(as.character(clindf$Cauris_Result))
clindf$Cauris_Result = plyr::revalue(clindf$Cauris_Result, c(
      "0"="Colonization Negative",
      "1"="Colonization Positive"))
clindf = subset(clindf, Survey_Period==1)
#fix the clinical data
Y1 = clindf$Cauris_Result
clindf = subset(clindf,  Unique_ptid != 32)
design = dplyr::select(clindf, c("Unique_ptid", "Site", "Survey_Period"))
clinSave = clindf
#clindf = subset(clindf, Unique_ptid != "32")
clindf$Unique_ptid = NULL
clindf$Site = NULL
clindf$Survey_Period = NULL
clindf$Cauris_Result = NULL

#scale the variables in the clinical table
clin_scales = scale(clindf)


#vst the taxa tables for each dataset independently
clrF = compositions::clr(data.frame(otu_table(its_filt)))
clrB = compositions::clr(data.frame(otu_table(bac_filt)))

#order the data frames so they are the same
clrB = clrB[rownames(clin_scales),]  
clrF = clrF[rownames(clin_scales),]  

### specify a design matrix
# in design we only need to mention the repeated measurements to split the one level variation
foo = subset_samples(its_filt, Survey_Period==1)
design1 = data.frame(sample_data(foo)) %>%
  dplyr::select("Unique_ptid")

design2 = data.frame(sample_data(foo)) %>%
  dplyr::select("SiteID")

### Within analysis
clinW <- mixOmics::withinVariation(X=clin_scales, design=design1)
clrBw <- mixOmics::withinVariation(X = clrB, design = design2)
clrFw <- mixOmics::withinVariation(X = clrF, design = design2)

### make table lists
Xa <- list(fungi=clrFw, bacteria=clrBw)

```




### Figure 2A:  Fungal data 

Panel A: Each panel encompasses samples for the specified body site with bars representing the relative abundance of taxa for each patient. The inner black curve represents the relative abundance of C. auris for each sample at that site. a) Relative abundance of top fungal genera at each body site surveyed for each individual. Genera included in the `Other` category include: Saccharomyces, Trichosporon, Trichophyton, Aspergillus.



```{r, fig.width=12, fig.height=5}
ordering =    c("Malassezia",
                "Malasseziales",    
                "Other",
                "Candida",
                "Candida utils",
                "Candida glabrata",
                "Candida duobushaemulonii",
                "Candida orthopsilosis",
                "Candida albicans",
                "Candida parapsilosis", 
                "Candida tropicalis",
                "Candida auris")

colours= c("Malassezia"="#77AADD",
            "Malasseziales" = "#77AADD",
             "Other"="#EE8866", 
             "Candida"= "#1B7837",
              "Candida utils"="#AA4499", #
             "Candida glabrata"="#AAAA00", #  
             "Candida duobushaemulonii"="#BBCC33", # 
             "Candida orthopsilosis"="#44BB99", #    
             "Candida albicans"="#99DDFF", #
            "Candida parapsilosis"="#993404", #
            "Candida tropicalis"="#EEDD88", #
            "Candida auris"="#E7D4E8")

label.vector = c( "Candida utils",
             "Candida glabrata", 
             "Candida duobushaemulonii",
             "Candida orthopsilosis",  
             "Candida albicans", 
            "Candida parapsilosis", 
            "Candida tropicalis", 
            "Candida auris")



```

### Figure 2

```{r}

show_candida_plot_by_rank <- function(phy, Site){
    map = data.frame(sample_data(phy)) %>%
      dplyr::select(., c("Fungal.Matcher", "percentAuris", "Cauris_Result"))
    rownames(map) = NULL
    myranks = map %>%
      arrange(., percentAuris) %>%
      mutate(., Rank=order(percentAuris))%>%
      dplyr::select(., c("Fungal.Matcher", "Rank"))
    myranks$Fungal.Matcher = str_replace_all(myranks$Fungal.Matcher, ";", ".")
    myranks$Fungal.Matcher = as.factor(as.character(myranks$Fungal.Matcher))
    mydf = phyToDf(phy, "Genus")
    df = plyr::join(mydf, myranks)
      df$Genus = str_remove_all(df$Genus, "g__")
      df$Genus = ifelse(!(df$Genus %in% c("Malassezia", "Candida", "Malasseziales")), "Other", df$Genus)
      df$Genus = ifelse(df$Genus=="Malasseziales", "Malassezia", df$Genus)
      df$Label = ifelse(df$Highest.Rank %in% label.vector, df$Highest.Rank, df$Genus)    
      df$Label=factor(df$Label, levels = ordering) 
      df = subset(df, Cauris_Result %in% c(0, 1))
    p = ggplot(df) +
            geom_col(aes(x=Rank, y=value, fill=Label), position="stack", width=1) + 
            geom_line(aes(x=Rank, y = percentAuris, group = 1)) + theme_classic() +
            ylab("") +  
            ylim(0, 1) +
            ggtitle(paste0(Site)) +
        theme(plot.title = element_text(size = 12),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
             scale_fill_manual(values = colours) 
    return(p)
    }

#set up the percent auris variable 
sample_data(its_match2)$percentAuris = sample_data(its_match2)$caurisReads/sample_sums(its_match2)
map = data.frame(sample_data(its_match2))

sample_data(bac_match) = sample_data(map)

myphy =  subset_samples(its_match2, SiteID=="An") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p1 = show_candida_plot_by_rank(phy=myphy, Site="Peri-Anus")  + 
             theme(legend.position = "none", 
                  text = element_text(size=14),
        axis.text.y = element_text(angle=0, hjust=1))  +
  ylab("Relative Abundance") 


myphy =  subset_samples(its_match2, SiteID=="Ax") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p2 = show_candida_plot_by_rank(phy=myphy, Site="Axilla") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())

myphy =  subset_samples(its_match2, SiteID=="Bu/To") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 

p3 = show_candida_plot_by_rank(phy=myphy, Site="Buccal Mucosa/Tongue") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())



myphy =  subset_samples(its_match2, SiteID=="N") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p4 = show_candida_plot_by_rank(phy=myphy, Site="Nares")  + 
             theme(legend.position = "none",
              axis.text.y = element_blank())

myphy =  subset_samples(its_match2, SiteID=="Ea") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p5 = show_candida_plot_by_rank(phy=myphy, Site="Ear canal")  + 
             theme(legend.position = "none",
              axis.text.y = element_blank())



myphy =  subset_samples(its_match2, SiteID=="Tc") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p6 = show_candida_plot_by_rank(phy=myphy, Site="Tracheostomy site")  + 
             theme(legend.position = "none",
              axis.text.y = element_blank())

myphy =  subset_samples(its_match2, SiteID=="Ic") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p7 = show_candida_plot_by_rank(phy=myphy, Site="Inguinal crease")  + 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(its_match2, SiteID=="Fg") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p8 = show_candida_plot_by_rank(phy=myphy, Site="Fingertips/Palms") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(its_match2, SiteID=="Tw") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p9 = show_candida_plot_by_rank(phy=myphy, Site="Toeweb") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(its_match2, SiteID=="Ne") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p10 = show_candida_plot_by_rank(phy=myphy, Site="Neck") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())

#### FIGURE LEGEND
mylegend = show_candida_plot_by_rank(phy=its_match2, Site="All")  + 
             theme(legend.position = "bottom") +guides(fill=guide_legend(ncol=10))

fungi_p = get_legend(mylegend)

```




```{r, fig.width=12}
### Make Figure 2a
Figure2_Legend <- cowplot::get_legend(mylegend)
Figure2 <- cowplot::plot_grid((p1 + theme(legend.position = "none")), 
                            (p2 + theme(legend.position = "none")), 
                            (p3 + theme(legend.position = "none")), 
                            (p4 + theme(legend.position = "none")),
                            (p5 + theme(legend.position = "none")),
                            (p6 + theme(legend.position = "none")), 
                            (p7 + theme(legend.position = "none")),
                            (p8 + theme(legend.position = "none")),
                            (p9 + theme(legend.position = "none")),                            
                            (p10 + theme(legend.position = "none")),                           
                          ncol = 10, hjust = -2.75, vjust = 1.5)
Figure2a <- cowplot::plot_grid(Figure2, Figure2_Legend, nrow = 2)
ggsave(Figure2a, file="~/Desktop/proctor_manuscript/Figure2/Figure2a.pdf",  width = 30, height = 6, units ="in",dpi = 300, device = "pdf")
ggsave(Figure2a, file="~/Desktop/proctor_manuscript/Figure2/Figure2a.eps",  width = 30, height = 6, units ="in",dpi = 300, device = "eps")

Figure2a
```

### Figure 2B: Plot the bacterial data 

Panel B: Each panel encompasses samples for the specified body site with bars representing the relative abundance of taxa for each patient. The inner black curve represents the relative abundance of C. auris for each sample at that site. b) Relative abundance of bacteria colored by Phylum reveals site-specific associations of C. auris with Proteobacteria.  



```{r, fig.height=3}

rank_samples_bygenus_and_plot <- function(phy, Site){
    map = data.frame(sample_data(phy)) %>%
      dplyr::select(., c("Fungal.Matcher", "percentAuris", "Cauris_Result"))
    rownames(map) = NULL
    myranks = map %>%
      arrange(., percentAuris) %>%
      mutate(., Rank=order(percentAuris))%>%
      dplyr::select(., c("Fungal.Matcher", "Rank"))
    myranks$Fungal.Matcher = as.factor(as.character(myranks$Fungal.Matcher))
    myranks$Fungal.Matcher = str_replace_all(myranks$Fungal.Matcher, ";", ".")
    
    mydf = phyToDf(myphy, "Phylum")
    df = plyr::join(mydf, myranks)
    df = subset(df, Cauris_Result %in% c(0, 1))
      p = ggplot(df) +
            geom_col(aes(x=Rank, y=value, fill=Phylum), position="stack", width=1)  +
            ylab("") +   scale_color_manual(values=ISU_secondary_palette) +
            scale_fill_manual(values=ISU_secondary_palette) + ylim(0, 1) +
            geom_line(aes(x=Rank, y = percentAuris, group = 1), size=1) + theme_classic()+
        theme(plot.title = element_text(size = 6),
              axis.title.x=element_blank(),
              axis.text.x=element_blank(),
              axis.ticks.x=element_blank())
    return(p)
    }


myphy =  subset_samples(bac_match, SiteID=="An") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p11 = rank_samples_bygenus_and_plot(phy=myphy, Site="Peri-anus")+ 
             theme(legend.position = "none", 
                  text = element_text(size=14),
        axis.text.y = element_text(angle=0, hjust=1))  +
  ylab("Relative Abundance") 



myphy =  subset_samples(bac_match, SiteID=="Ax") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p12 = rank_samples_bygenus_and_plot(phy=myphy, Site="Axilla") + 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(bac_match, SiteID=="Bu/To") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p13 = rank_samples_bygenus_and_plot(phy=myphy, Site="Buccal Mucosa / Tongue")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(bac_match, SiteID=="N") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p14 = rank_samples_bygenus_and_plot(phy=myphy, Site="Nares")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())

myphy =  subset_samples(bac_match, SiteID=="Ea") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p15 = rank_samples_bygenus_and_plot(phy=myphy, Site="Ear canal")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())



myphy =  subset_samples(bac_match, SiteID=="Tc") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p16 = rank_samples_bygenus_and_plot(phy=myphy, Site="Tracheostomy site")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())

myphy =  subset_samples(bac_match, SiteID=="Ic") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p17 = rank_samples_bygenus_and_plot(phy=myphy, Site="Inguinal crease")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(bac_match, SiteID=="Fg") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p18 = rank_samples_bygenus_and_plot(phy=myphy, Site="Fingertips/Palms")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(bac_match, SiteID=="Tw") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p19 = rank_samples_bygenus_and_plot(phy=myphy, Site="Toeweb")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())


myphy =  subset_samples(bac_match, SiteID=="Ne") %>%
      prune_samples(sample_sums(.) > 0, .) %>%
      prune_taxa(taxa_sums(.) > 0, .) 
p20 = rank_samples_bygenus_and_plot(phy=myphy, Site="Neck")+ 
             theme(legend.position = "none",
              axis.text.y = element_blank())

mylegend = rank_samples_bygenus_and_plot(phy=bac_match, Site="Peri-Anus") +
             theme(legend.position = "bottom") +guides(fill=guide_legend(ncol=10))
```



```{r, fig.width=12}
### Make Figure 2b
Figure2_Legend <- cowplot::get_legend(mylegend)
Figure2 <- cowplot::plot_grid((p11 + theme(legend.position = "none")), 
                            (p12 + theme(legend.position = "none")), 
                            (p13 + theme(legend.position = "none")),
                            (p14 + theme(legend.position = "none")),
                            (p15 + theme(legend.position = "none")),
                            (p16 + theme(legend.position = "none")), 
                            (p17 + theme(legend.position = "none")),
                            (p18 + theme(legend.position = "none")),
                            (p19 + theme(legend.position = "none")),                            
                            (p20 + theme(legend.position = "none")),                           
                          ncol = 10, hjust = -2.75, vjust = 1.5)
Figure2b <- cowplot::plot_grid(Figure2, Figure2_Legend, nrow = 2)
ggsave(Figure2b, file="~/Desktop/proctor_manuscript/Figure2/Figure2b.eps",  width = 30, height = 6, units ="in",dpi = 300, device = "eps")
ggsave(Figure2b, file="~/Desktop/proctor_manuscript/Figure2/Figure2b.pdf",  width = 30, height = 6, units ="in",dpi = 300, device = "pdf")
Figure2b

```


### Now let's look at variables int he bacterial table that segregate with colonization status
```{r, fig.height=5, fig.width=5}
bacterial.splsda <- mixOmics::splsda(clrB, Y1, ncomp=2, near.zero.var=TRUE) # 1 Run the method
mixOmics::plotIndiv(bacterial.splsda, ellipse = TRUE, star=TRUE, ind.names = FALSE)                          # 2 Plot the samples
```

```{r}
setEPS()
pdf("~/Desktop/proctor_manuscript/Figure2/Figure2c.pdf")

par(mfrow=c(1,1)) 
mixOmics::plotIndiv(bacterial.splsda, ellipse = TRUE, star=TRUE, ind.names = FALSE)                          # 2 Plot the samples
dev.off()
```


```{r}
tax1 = data.frame(bac_filt@tax_table@.Data) %>%
  dplyr::select(., c("Kingdom", "Class", "Order", "Family", "Genus", "Highest.Rank"))
tax1$Highest.Rank = str_replace_all(tax1$Highest.Rank, " ", ".")
loadings.16s = data.frame(bacterial.splsda$loadings$X)
loadings.16s$Highest.Rank= rownames(loadings.16s)
loadings.16s$Highest.Rank = str_replace_all(loadings.16s$Highest.Rank, " ", ".")
loadings.16s = plyr::join(loadings.16s, tax1)

loadings.16s = loadings.16s[with(loadings.16s, order(comp1)),]
ordering16s = as.vector(loadings.16s$Highest.Rank)
loadings.16s$Highest.Rank  <- factor(loadings.16s$Highest.Rank, levels = ordering16s)
loadings.16s$group = ifelse(loadings.16s$comp1 < 0, "Culture Negative", "Culture Positive")


myfont <- element_text(face = "italic",size = 12)

Figure2d = ggplot() +
  geom_col(data=loadings.16s, aes(Highest.Rank, comp1, fill=group)) +
  coord_flip()+ scale_fill_manual(values=c("#6699FF", "#FF9933"), name="Group")+
  theme_classic() +
  ylab("Component 1")+ xlab("")+ theme(axis.text.y = myfont)


Figure2d

ggsave(Figure2d, file="~/Desktop/proctor_manuscript/Figure2/Figure2d.eps",  width = 6, height = 6, units ="in",dpi = 300, device = "eps")

```

### Let's look at thee performance of the model 
```{r}
#Performance 
set.seed(123) # for reproducibility 
MyPerf.selected.diablo <- mixOmics::perf(bacterial.splsda, validation = 'Mfold', folds = 2, 
                   nrepeat = 100, 
                   dist = c('centroids.dist', "max.dist", "mahalanobis.dist")) #,                    design=design3

bacterial.splsda$choice.ncomp
bacterial.splsda$MajorityVote.error.rate
auroc(bacterial.splsda)
```


#### Let's build a multiple table model


Select variables in the bacterial table
```{r, fig.height=4, fig.width=4}
bacterial.splsda <- splsda(clrB, Y1, ncomp=8, near.zero.var=TRUE) # 1 Run the method
plotIndiv(bacterial.splsda, ellipse = TRUE, star=TRUE, ind.names = FALSE)                          # 2 Plot the samples
plotVar(bacterial.splsda, cutoff=0.4, var.names = FALSE)
plotLoadings(bacterial.splsda)
keep.bacterial.taxa = selectVar(bacterial.splsda, comp=1)$name   
```

Select variables in the fungal table
```{r, fig.height=4, fig.width=4}
fungal.splsda <- splsda(clrF, Y1, ncomp=8, near.zero.var=TRUE) # 1 Run the method
plotIndiv(fungal.splsda, ellipse = TRUE, star=TRUE, ind.names = FALSE)                          # 2 Plot the samples
plotVar(fungal.splsda, cutoff=0.4, var.names = TRUE)
plotVar(fungal.splsda, cutoff=0.4, var.names = FALSE)
keep.fungal.taxa = selectVar(fungal.splsda, comp=1)$name   
```

subset the taxa on the selected variables
```{r}
#select the taxa
clrB2 <- clrB[, keep.bacterial.taxa]
clrF2 <- clrF[, keep.fungal.taxa]

#order the data frames so they are the same
clrB2 = clrB2[rownames(clin_scales),]  
clrF2 = clrF2[rownames(clin_scales),]  


### Within analysis
clinW <- withinVariation(X=clin_scales, design=design1)
clrBw <- withinVariation(X = clrB2, design = design2)
clrFw <- withinVariation(X = clrF2, design = design2)

### make table lists
Xa <- list(clinical=clin_scales, fungi=clrFw, bacteria=clrBw)
Xa <- list( fungi=clrFw, bacteria=clrBw)
```


### Run the first model; use the first time point; account for within site variation

```{r, fig.height=4, fig.width=8}
set.seed(861)
MyResult.diablo <- block.splsda(Xa, Y1,    scale=TRUE,   near.zero.var=TRUE, scheme="centroid", ncomp = 4)
plotIndiv(MyResult.diablo, 
          ind.names = FALSE, 
          ellipse = TRUE,
          star=TRUE,
          legend=TRUE,
          abline = TRUE)

```


```{r, fig.height=4, fig.width=4}
plotVar(MyResult.diablo,
      comp = c(1, 2),
      cutoff = 0.25,
      rad.in = 0.5,
      cex=c(4, 4),
      var.names = FALSE,
      style="ggplot2",
      overlap = TRUE,
      axes.box = "all",
      label.axes.box = "both")
```


```{r, fig.height=4, fig.width=4}
plotVar(MyResult.diablo,
      comp = 1:2,
      cutoff = 0.25,
      rad.in = 0.5,
      cex=c(3, 3),
      var.names = TRUE,
      style="ggplot2",
      overlap = TRUE,
      axes.box = "all",
      label.axes.box = "both")

```


```{r, fig.height=8}
tax1 = data.frame(bac_match@tax_table@.Data) %>%
  dplyr::select(., c("Kingdom", "Class", "Order", "Family", "Genus", "Highest.Rank"))
tax1$Highest.Rank = str_replace_all(tax1$Highest.Rank, " ", ".")
loadings.16s = data.frame(MyResult.diablo$loadings$bacteria)
loadings.16s$Highest.Rank= rownames(loadings.16s)
loadings.16s$Highest.Rank = str_replace_all(loadings.16s$Highest.Rank, " ", ".")
loadings.16s = plyr::join(loadings.16s, tax1)

loadings.16s = loadings.16s[with(loadings.16s, order(comp1)),]
ordering16s = as.vector(loadings.16s$Highest.Rank)
loadings.16s$Highest.Rank  <- factor(loadings.16s$Highest.Rank, levels = ordering16s)


tax1 = data.frame(its_match@tax_table@.Data) %>%
  dplyr::select(., c("Kingdom", "Class", "Order", "Family", "Genus", "Highest.Rank"))
tax1$Highest.Rank = str_replace_all(tax1$Highest.Rank, " ", ".")
loadings.its = data.frame(MyResult.diablo$loadings$fungi)
loadings.its$Highest.Rank= rownames(loadings.its)
loadings.its$Highest.Rank = str_replace_all(loadings.its$Highest.Rank, " ", ".")
loadings.its = plyr::join(loadings.its, tax1)
loadings.its = loadings.its[with(loadings.its, order(comp1)),]
orderingITS = as.vector(loadings.its$Highest.Rank)
loadings.its$Highest.Rank  <- factor(loadings.its$Highest.Rank, levels = orderingITS)

  
### Coordinate 1
p1 = ggplot() +
  geom_col(data=loadings.its, aes(Highest.Rank, comp1)) +
  coord_flip()
  
p2 = ggplot() +
  geom_col(data=loadings.16s, aes(Highest.Rank, comp1)) +
  coord_flip()

### Coordinate 2
loadings.16s = loadings.16s[with(loadings.16s, order(comp2)),]
ordering16s = as.vector(loadings.16s$Highest.Rank)
loadings.16s$Highest.Rank  <- factor(loadings.16s$Highest.Rank, levels = ordering16s)

loadings.its = loadings.its[with(loadings.its, order(comp2)),]
orderingITS = as.vector(loadings.its$Highest.Rank)
loadings.its$Highest.Rank  <- factor(loadings.its$Highest.Rank, levels = orderingITS)



p3 = ggplot() +
  geom_col(data=loadings.its, aes(Highest.Rank, comp2)) +
  coord_flip()
  
p4 = ggplot() +
  geom_col(data=loadings.16s, aes(Highest.Rank, comp2)) +
  coord_flip()

grid.arrange(p1, p2, p3, p4, ncol=2)
```


#### Let's look at explicit correlations
Since the multiple table model threw an error saying the direction of the correlation may not be accurately represented in the model let's just look at straight up correlations between each taxa's distribution and the distribution of Candida auris across samples. We will use a linear mixed effects model to account for multiple measures within an indidivudal

`### Mixed effects sequencing data - The model: abundance ~ cauris chg_conc + SiteID + (1 | Unique_ptid) +ϵ

Merge the its and bacterial tables
```{r}
#get rid of the tree - fungal
otus = otu_table(its_match)
map = sample_data(its_match)
tax = its_match@tax_table@.Data
its_match = merge_phyloseq(otus, map, tax_table(tax))


#subset the its table to eliminate noisy taxa; otherwise model fails
library(DESeq2);library(genefilter)
filtergroup = genefilter::filterfun(genefilter::kOverA(k=20, A=10)) #k = number of samples; A = abundance
#filter  taxa 
        filtPhy = filter_taxa(its_match, filtergroup, prune=TRUE) 
        filtPhy = prune_taxa(taxa_sums(filtPhy) > 0, filtPhy) 
        filt_its = subset_samples(filtPhy, Unique_ptid != 32)

#generate a combined fungal/bacterial table so we can adjust pvalues in the regression appropriately
phy = merge_phyloseq(bac_match, filt_its)      
```


```{r}
set.seed(78927)
phy = transform_sample_counts(phy, function(x) compositions::clr(x)) 

#make a map for the regression
map = data.frame(sample_data(phy)) %>%
  dplyr::select(., c("sqrt_cauris", "Unique_ptid", "SiteID"))

#convert site and survey period to numeric
map$SiteID = as.numeric(factor(map$SiteID))
map = data.frame(scale(map))

#get the otu table of the centered log ratio table
otus = data.frame(otu_table(phy))

attach(map)
all=data.frame(cbind(otus, map))

#set up empty lists
mod <- list()
out <- list()
adjp <- list()

#https://stackoverflow.com/questions/57590176/adjust-p-values-obtained-with-lmertestlmer-for-multiple-comparisons
adjMC <- function( model_name ) {
  model_glht <- glht(model_name)
  model_MCadj <- summary(model_glht, test = adjusted('holm')) # Bonferroni-Holm 
  return(model_MCadj)
}

library(multcomp)
for(i in names(otus)[-1]){
  mod[[i]] <- lmerTest::lmer(get(i) ~ sqrt_cauris   + SiteID +
                               (1 | Unique_ptid  ), 
      data = all)
  adjp[[i]]  = adjMC(mod[[i]])
  out[[i]] = broom.mixed::tidy(adjp[[i]], conf.int = TRUE, .name_repair = "unique")
}
tax = data.frame(phy@tax_table@.Data) 
out = out %>% map_dfr(~ .x %>% as_tibble(), .id = "Highest.Rank") 
out$Highest.Rank = str_replace_all(out$Highest.Rank, "([.])", " ")

df = data.frame(out) %>%
  plyr::join(tax) %>%
  subset(., contrast=="sqrt_cauris")  

#make a volcano plot
library(RColorBrewer)
pal <- brewer.pal(n = 4, name = 'Set1')
df$adj.p.value = as.numeric(as.character(df$adj.p.value))
my.annotation = subset(df, adj.p.value  < 0.05 & estimate > 0.1 | estimate <= -0.1)
anno2 = subset(df, Highest.Rank %in% c("Acinetobacter baumannii", "Pseudomonas aeruginosa"))
my.annotation = data.frame(rbind(my.annotation, anno2))
df$Highest.Rank = as.character(as.factor(df$Highest.Rank))
df$neglog = -log10(df$adj.p.value)
df$neglog  = ifelse(df$neglog=="Inf", 12, df$neglog)
      
## Create a column to indicate which genes to label
df$species.label = ifelse(df$Highest.Rank %in% my.annotation$Highest.Rank, "TRUE", "FALSE")
myannotations = subset(df, species.label==TRUE)
```

let's see a volcano plot of significant taxa
```{r}
Figure3C = ggplot(df) +
  geom_point(aes(x = estimate, y = neglog), alpha=0.5) +
  geom_text_repel(data=myannotations, aes(x = estimate, y = neglog, label = Highest.Rank),size=4) +
  xlab("Regression Coefficient") + 
  ylab("-log10 adjusted p-value")  +
  xlim(-1.2, 1.2)  +
  geom_vline(xintercept = -0.1, linetype='dashed', color="gray") + 
  geom_vline(xintercept = 0.1, linetype='dashed', color="gray") +
  geom_hline(yintercept = -log(0.05),  linetype='dashed', color="gray") + 
  theme_classic() +
  scale_color_manual(values=c("#377EB8" ,"#4DAF4A"))+
    theme(text = element_text(size=14),
        axis.text.x = element_text(angle=0, hjust=1)) + theme(legend.position = "none") 

Figure3C

```


print the table
```{r}
myannotations = subset(df, adj.p.value  < 0.05 )
myannotations = myannotations[with(myannotations, order(estimate)),]

myannotations$upper = myannotations$std.error + myannotations$estimate
myannotations$lower = myannotations$estimate- myannotations$std.error
ordering = as.vector(myannotations$Highest.Rank)
myannotations$Highest.Rank  <- factor(myannotations$Highest.Rank, levels = ordering)
p = ggplot(myannotations, aes(Highest.Rank, estimate, color=Phylum)) + geom_point() +
        geom_errorbar(aes(ymin=upper, ymax=lower), width=.2,
                 position=position_dodge(.9)) + coord_flip() + 
  theme_classic()

p
```
